---
title: "STAT 545A Homework 4"
author: "Junbin ZHANG"
date: "Sept 25, 2018"
output:
  html_document:
    keep_md: true
    toc: true
---

# Bring rectangular data in
```{r}
## load gapminder
library(gapminder)
## load tidyverse
library(tidyverse)
```

# Data Reshaping Prompts (and relationship to aggregation)

## Activity #1

**Make your own cheatsheet.**

In this section, we try to explore the following functions in `tidyr`:

- `gather()`: converts wide data to longer format.
- `spread()`: converts long data to wider format.
- `unite()`: combines two or more columns into a single column.
- `separate()`: splits one column into two or more columns.

We are going to use `mtcars` and `gapminder` as the databases.

```{r}
# show a preview of mtcars
head(mtcars) %>% 
  knitr::kable()
```

We give the first column a name ("car") so it is better to manipulate. In fact, the first column in the above table is the `rownames()` of the table.

```{r}
# add a new column and assign rownames() to it
mtcars$car <- rownames(mtcars)
# reshape the data frame
mtcars_with_car <- mtcars[, c(12, 1:11)]
# show a preview of modified mtcars
head(mtcars_with_car) %>% 
  knitr::kable()
```


### `gather()`

#### Format of `gather()`

`gather(data, key, value, ..., na.rm = FALSE, convert = FALSE)`

Here `...` is the specification of the columns to gather.

#### Demos

`gather()` tries to create key-value pairs for a given data frame. In the first demos, we use `car` as the index, and different attribute as keys and show their corresponding values.

```{r}
gather_attributes <- mtcars_with_car %>% 
  # list all different attributes in column "attribute", use "car" as index (so it is not included in the reshape)
  gather(attribute, value, -car)

# show previews 
head(gather_attributes) %>% 
  knitr::kable()

tail(gather_attributes) %>% 
  knitr::kable()
```

We can also gather some columns instead of all.

```{r}
gather_some_attributes <- mtcars_with_car %>% 
  # gather only mpg to gear
  gather(attribute, value, mpg:gear)

# show previews
head(gather_some_attributes) %>% 
  knitr::kable()

tail(gather_some_attributes) %>% 
  knitr::kable()
```

### `spread()`

#### Format of `spread()`

`spread(data, key, value, fill = NA, convert = FALSE, drop = TRUE)`

#### Demos

`spread()` can be considered as a reversed operation of `gather()`.

```{r}
# spread attribute and value
mtcars_spread <- gather_attributes %>% 
  spread(attribute, value)

# show a preview
head(mtcars_spread) %>% 
  knitr::kable()
```

### `unite()`

#### Format of `unite()`

`unite(data, col, ..., sep = "_", remove = TRUE)`

Here `...` represents the columns to unite and `col` represents the column to add.

#### Demos

In this demo, we try to unite `pop` and `gdpPercap` into the same column in `gapminder`. Notice it just shows the new column as a multiplication of `pop` and `gdpPercap`, but no calculation will be performed.

```{r}
# unite to have overall_gdp
unite_gdp <- gapminder %>% 
  unite(overall_gdp, pop, gdpPercap, sep = "*")

# show a preview
head(unite_gdp) %>% 
  knitr::kable()
```


### `separate()`

#### Format of `separate()`

`separate(data, col, into, sep = "[^[:alnum:]]+", remove = TRUE, convert = FALSE, extra = "warn", fill = "warn", ...)`

#### Demos

`separate()` can be considered as a reversed operation of `unite()`.

```{r}
# separate into original columns. Notice that we need to use two backslashes as escape character.
sep_gdp <- unite_gdp %>% 
  separate(overall_gdp, c("pop", "gdpPercap"), sep="\\*")

# show a preview
head(sep_gdp) %>% 
  knitr::kable()
```

## Activity #2

**Make a tibble with one row per year and columns for life expectancy for two or more countries.**

Let's select countries in Oceania.

```{r}
lifeExp_Oceania <- gapminder %>% 
  # filter countries in Oceania
  filter(continent == "Oceania") %>% 
  # select necessary columns, make year as the first column
  select(year, country, lifeExp) %>% 
  # spread country to columns, with lifeExp as values
  spread(country, lifeExp)

# display the table
lifeExp_Oceania %>% 
  knitr::kable()
```

In Oceania we only have two countries. Then we make a scatterplot of the new tibble. First, we show the trends of lifeExp over years in these two countries.

```{r}
# fix names with spaces
names(lifeExp_Oceania) <- str_replace_all(names(lifeExp_Oceania), " ", "_")

lifeExp_Oceania %>% 
  ggplot() +
  # make a line plot with points for Australia
  geom_line(aes(x=year, y=Australia, color="Australia")) +
  geom_point(aes(x=year, y=Australia, color="Australia")) +
  # make a line plot with points for New_Zealand
  geom_line(aes(x=year, y=New_Zealand, color="New Zealand")) +
  geom_point(aes(x=year, y=New_Zealand, color="New Zealand")) +
  # change color legent
  scale_color_discrete("country")
```

Let's also compare them side-by-side.

```{r}
lifeExp_Oceania %>% 
  # Auatralia as x axis and New_Zealand as y axis, color as year
  ggplot(aes(x=Australia, y=New_Zealand, color=factor(year))) +
  # make it a scatterplot
  geom_point() +
  # try to draw a regression curve
  geom_smooth(method="lm", se=FALSE, color="black")
```

We can learn from the figures that, after around 1973, the lifeExp in Australia becomes larger than New_Zealand.

## Activity #3

**Compute some measure of life expectancy (mean? median? min? max?) for all possible combinations of continent and year. Reshape that to have one row per year and one variable for each continent. Or the other way around: one row per continent and one variable per year.**

Let's first compute the means of life expectancy for every possible combinations of continent and year.

```{r}
mean_lifeExp <- gapminder %>% 
  # group by continent and year
  group_by(continent, year) %>%
  # calculate the means
  summarise(
    mean_lifeExp = mean(lifeExp)
  )

# show the table
mean_lifeExp %>% 
  knitr::kable()
```

Then we reshape it to one row per year and one variable for each continent.

```{r}
# use spread to reshape continent as columns
reshape_mean_lifeExp <- mean_lifeExp %>%
  spread(continent, mean_lifeExp)

# show the table
reshape_mean_lifeExp %>% 
  knitr::kable()
```

The reshaped tibble is better, if you only want to have data of one continent. For example, we try to get data of Africa and show it as line plot.

```{r}
# original tibble
mean_lifeExp %>% 
  # need to filter continent as Africa
  filter(continent == "Africa") %>% 
  # year as x axis and mean_lifeExp as y axis
  ggplot(aes(x=year, y=mean_lifeExp)) +
  # make it a line plot with point
  geom_line() +
  geom_point()

reshape_mean_lifeExp %>% 
  # year as x axis and Africa as y axis
  ggplot(aes(x=year, y=Africa)) +
  # make it a line plot with point
  geom_line() +
  geom_point()
```

The two plots are basically the same, except:

- Using the origianl tibble we need to filter data;
- The default label of y axis in the second figure is "Africa", and we do not need to change it.

## Activity #4

**In [Window functions](http://stat545.com/block010_dplyr-end-single-table.html#window-functions), we formed a tibble with 24 rows: 2 per year, giving the country with both the lowest and highest life expectancy (in Asia). Take that table (or a similar one for all continents) and reshape it so you have one row per year or per year * continent combination.**

First we get the table from the reference.

```{r}
# this piece of code is directly from the reference, so no commend is added
my_gap <- gapminder %>%
  filter(continent == "Asia") %>%
  select(year, country, lifeExp) %>%
  group_by(year) %>%
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>% 
  arrange(year) %>%
  print(n = Inf)
```

We reshape it into one row per year.

```{r}
# unite country and lifeExp
my_gap_unite <- my_gap %>% 
  unite(country_lifeExp, country, lifeExp, sep=",")

# get maximum records for each year (even rows)
my_gap_max <- my_gap_unite %>% 
  filter(row_number() %% 2 == 0)

# get minimum records for each year (odd rows)
my_gap_min <- my_gap_unite %>% 
  filter(row_number() %% 2 == 1)

# combine data together
my_gap_final <- data.frame(
  year = my_gap_max$year,
  max_country_lifeExp = my_gap_max$country_lifeExp,
  min_country_lifeExp = my_gap_min$country_lifeExp
)

# show the table
my_gap_final %>% 
  knitr::kable()
```

## Activity #5

**Make up a similar set of exercises for yourself, in the abstract or (even better) using Gapminder or other data, and solve them.**

According to the [data manipulation sample](https://gist.github.com/aammd/11386424), some of the functions have been already covered in the previous assignemtns:

- Group using `group_by()` and show statistics using `summarize()`.
- Subset using `filter()`.
- Sort using `arragnge()`.

We are going to explore the following functions in the next section, so we can also skip them right now:

- Join using `left_join()` and `merge()`.

Therefore, we focus on the following functions:

- Split using `dcast()`.
- Stack using `melt()`.
- Transpose using `t()`.

Some of these functions belong to package `reshape2`, and assuming it is already installed.

```{r}
# load reshape2
library(reshape2)
```


### Split

Here we show that `dcast()` is actually similar to `spread()`. The following example is retrieved form Activity #4.

```{r}
split <- my_gap %>% 
  # year as rows and country as columns, fill lifeExp as values
  dcast(year~country, value.var="lifeExp")

# display the table
split %>% 
  knitr::kable()
```

### Stack

Here we show that `melt()` is actually similar to `group()`. The following example is retrieved from Activity #1.

```{r}
melt_attributes <- mtcars_with_car %>% 
  # melt and use car as index
  melt(id="car")

# show previews 
head(melt_attributes) %>% 
  knitr::kable()

tail(melt_attributes) %>% 
  knitr::kable()
```


### Transpose

Let's try to transpose `mtcars`.

```{r}
# remove car
mtcars_ori <- mtcars %>% 
  select(-car)

# transpose data frame
mtcars_transpose <- data.frame(t(mtcars_ori))

# display the 
mtcars_transpose %>% 
  knitr::kable()
```

# Join Prompts (join, merge, look up)

## Activity #1

**Create a second data frame, complementary to Gapminder. Join this with (part of) Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins.**

According to [Data from Gapminder](https://cran.r-project.org/web/packages/gapminder/gapminder.pdf), there is a data frame called `country_codes` in package `gapminder`, which is suitable in our example.

```{r}
# show a preview of country_code
head(country_codes) %>% 
  knitr::kable()
```

`country_codes` contains rows for all country, so somehow for every row in the full `gapminder` data frame, we can find at least one matching records in `country_codes`. To make it more intersting, we trim `gapminder` to include only countries in "Americas". We also try to select a single year, say 1952.

```{r}
gapminder_Americas <- gapminder %>% 
  # filter continent as Americas
  filter(continent == "Americas", year == 1952)

# show a preview
head(gapminder_Americas) %>% 
  knitr::kable()
```


We explore the following join functions:

- `inner_join()`: return all rows from x where there are matching values in y, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.
- `semi_join()`: return all rows from x where there are matching values in y, keeping just columns from x. A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join.
- `left_join()`: return all rows from x, and all columns from x and y. If there are multiple matches between x and y, all combination of the matches are returned. This is a mutating join.
- `anti_join()`: return all rows from x where there are not matching values in y, keeping just columns from x. This is a filtering join.
- `full_join()`: return all rows and all columns from both x and y. Where there are not matching values, returns NA for the one missing. This is a mutating join.

### `inner_join()`

In the following two examples, we change the order of `x` and `y`.

```{r}
inner_join(x=gapminder_Americas, y=country_codes) %>% 
  knitr::kable()
```

```{r}
inner_join(x=country_codes, y=gapminder_Americas) %>% 
  knitr::kable()
```

After `inner_join()`, we get rows with countries only exist in `gapminder_Americas`, but for each row, we have columns both from `gapminder_Americas` and `country_codes`. Changing the order of `x` and `y` does not affect the results, since `inner_join()` tries to find overlapping countries in two data frames.

### `semi_join()`

In the following two examples, we change the order of `x` and `y`.

```{r}
semi_join(x=gapminder_Americas, y=country_codes) %>% 
  knitr::kable()
```

```{r}
semi_join(x=country_codes, y=gapminder_Americas) %>% 
  knitr::kable()
```

Compare to `inner_join()`, `semi_join()` also returns overlapping countries only as rows. However, the columns depend on which data frame is provided as `x`.

### `left_join()`

In the following two examples, we change the order of `x` and `y`.

```{r}
left_join(x=gapminder_Americas, y=country_codes) %>% 
  knitr::kable()
```

```{r}
left_join(x=country_codes, y=gapminder_Americas)  %>% 
  knitr::kable()
```

Using `left_join()`, rows depend on what countries we have in `x`. All columns in `x` and `y` are used to form the final table. However, since we cannot find all countries in `gapminder_Americas`, for those countries only exist in `country_codes`, the values are `NA` for columns in `gapminder_Americas`.

### `anti_join()`

In the following two examples, we change the order of `x` and `y`.

```{r}
anti_join(x=gapminder_Americas, y=country_codes) %>% 
  knitr::kable()
```

```{r}
anti_join(x=country_codes, y=gapminder_Americas)  %>% 
  knitr::kable()
```

`anti_join()` is interesting that it shows rows in `x` that cannot be found a matching value in `y`, and columns to shown depend on `x`:

- Since one can find all country in `gapminder_Americas` in `country_codes`, the first table is empty.
- The second table actually shows countries that do not exist in `gapminder_Americas`.

### `full_join()`

In the following two examples, we change the order of `x` and `y`.

```{r}
full_join(x=gapminder_Americas, y=country_codes) %>% 
  knitr::kable()
```

```{r}
full_join(x=country_codes, y=gapminder_Americas)  %>% 
  knitr::kable()
```

It is expected that the above two tables generated by `full_join()` have the same results, since this function shows all rows and columns from both `x` and `y`. Due to the data frames we choose (countries in `gapminder_Americas` is actually a subset of `country_codes`), the resultant tables are with the identical data as running `left_join(x=country_codes, y=gapminder_Americas)`.

## Activity #3

### `merge()`

In the following two examples, we change the order of `x` and `y`.

```{r}
merge(x=gapminder_Americas, y=country_codes) %>% 
  knitr::kable()
```

```{r}
merge(x=country_codes, y=gapminder_Americas) %>% 
  knitr::kable()
```

From the results, by default, we can conclude that `merge()` is the same as `inner_join()`.

Actually it is also possible to use `merge()` to implement `full_join()` and `left_join()`.

#### Implement `full_join()` using `merge()`

```{r}
merge(x=gapminder_Americas, y=country_codes, all=TRUE) %>% 
  knitr::kable()
```

#### Implement `left_join()` using `merge()`

Notice that the order of `x` and `y` is important in this example.

```{r}
merge(x=gapminder_Americas, y=country_codes, all.x=TRUE) %>% 
  knitr::kable()
```

### `match()`

```{r}
match(gapminder_Americas$country, country_codes$country) %>% 
  knitr::kable()
```

For the example we can know, there are some disadvantages using `match()` compared to `merge()` or join functions:

- It only works for two vectors.
- It only returns indeces of matching items.

# References
- [Data manipulation with tidyr](https://www.r-bloggers.com/data-manipulation-with-tidyr/)
- [Cheatsheet for dplyr join functions](http://stat545.com/bit001_dplyr-cheatsheet.html)