---
title: "STAT 545A Homework 4"
author: "Junbin ZHANG"
date: "Sept 25, 2018"
output:
  html_document:
    keep_md: true
---

# Bring rectangular data in
```{r}
## load gapminder
library(gapminder)
## load tidyverse
library(tidyverse)
```

# Data Reshaping Prompts (and relationship to aggregation)

## Activity #1

In this section, we try to explore the following functions in `tidyr`:
- `gather()`: converts wide data to longer format.
- `spread()`: converts long data to wider format.
- `unite()`: combines two or more columns into a single column.
- `separate()`: splits one column into two or more columns.

We are going to use `mtcars` and `gapminder` as the databases.

```{r}
# show a preview of mtcars
head(mtcars) %>% 
  knitr::kable()
```

We give the first column a name ("car") so it is better to manipulate. In fact, the first column in the above table is the `rownames()` of the table.

```{r}
# add a new column and assign rownames() to it
mtcars$car <- rownames(mtcars)
# reshape the data frame
mtcars_with_car <- mtcars[, c(12, 1:11)]
# show a preview of modified mtcars
head(mtcars_with_car) %>% 
  knitr::kable()
```


### `gather()`

#### Format of `gather()`

`gather(data, key, value, ..., na.rm = FALSE, convert = FALSE)`

Here `...` is the specification of the columns to gather.

#### Demos

`gather()` tries to create key-value pairs for a given data frame. In the first demos, we use `car` as the index, and different attribute as keys and show their corresponding values.

```{r}
gather_attributes <- mtcars_with_car %>% 
  # list all different attributes in column "attribute", use "car" as index (so it is not included in the reshape)
  gather(attribute, value, -car)

# show previews 
head(gather_attributes) %>% 
  knitr::kable()

tail(gather_attributes) %>% 
  knitr::kable()
```

We can also gather some columns instead of all.

```{r}
gather_some_attributes <- mtcars_with_car %>% 
  #gather only mpg to gear
  gather(attribute, value, mpg:gear)

# show previews
head(gather_some_attributes) %>% 
  knitr::kable()

tail(gather_some_attributes) %>% 
  knitr::kable()
```

### `spread()`

#### Format of `spread()`

`spread(data, key, value, fill = NA, convert = FALSE, drop = TRUE)`

#### Demos

`spread()` can be considered as a reversed operation of `gather()`.

```{r}
# spread attribute and value
mtcars_spread <- gather_attributes %>% 
  spread(attribute, value)

# show a preview
head(mtcars_spread) %>% 
  knitr::kable()
```

### `unite()`

#### Format of `unite()`

`unite(data, col, ..., sep = "_", remove = TRUE)`

Here `...` represents the columns to unite and `col` represents the column to add.

#### Demos

In this demo, we try to unite `pop` and `gdpPercap` into the same column in `gapminder`. Notice it just shows the new column as a multiplication of `pop` and `gdpPercap`, but no calculation will be performed.

```{r}
# unite to have overall_gdp
unite_gdp <- gapminder %>% 
  unite(overall_gdp, pop, gdpPercap, sep = "*")

# show a preview
head(unite_gdp) %>% 
  knitr::kable()
```


### `separate()`

#### Format of `separate()`

`separate(data, col, into, sep = "[^[:alnum:]]+", remove = TRUE, convert = FALSE, extra = "warn", fill = "warn", ...)`

#### Demos

`separate()` can be considered as a reversed operation of `unite()`.

```{r}
# separate into original columns. Notice that we need to use two backslashes as escape character.
sep_gdp <- unite_gdp %>% 
  separate(overall_gdp, c("pop", "gdpPercap"), sep="\\*")

# show a preview
head(sep_gdp) %>% 
  knitr::kable()
```

## Activity #2

## Activity #3

## Activity #4

## Activity #5

# Join Prompts (join, merge, look up)

## Activity #1

## Activity #2

## Activity #3

# References
- [Data manipulation with tidyr](https://www.r-bloggers.com/data-manipulation-with-tidyr/)